Title

### 2024-06-22 paper/main.tex
Updates

In this section, we develop a (conditional) measure of association for mixed
data as a generalization of partial correlation by combining a mixed data
residualization approach with canonical correlations. Assuming that we need to
compute the association, $ \rho(X, Y; Z) $ between variables $ X $ and $ Y $
conditioned on $ \bm{Z} $. Similar to partial correlations, we first start by
computing the residuals for both $ X $ and $ Y $ using $ \bm{Z} $ as the
predictor variables. When $ \bm{Z} = \emptyset $ we can use a constant value
for $ \bm{Z} $. We use the mixed data residualization approach presented in
Ankan and Textor 2023.

\subsection{Mixed Data Residualization}
Residualization for mixed data is not as straight forwards as for continuous
data. We use the residualization approach proposed in Ankan 2023. Given a variable
$ X $ conditioned on $ \bm{Z} $, we compute the residuals as follows:

If $ X $ is continuous, we simply take the difference between the true and the 
predicted value.

If $ X $ is ordinal, we take the LS-Residuals

If $ X $ is categorical, we one hot encode the variable.

As a result of this residualization we can get a matrix of continuous values for 
the residuals. We need a measure of association for this set of variables.

\subsection{Canonical Correlation}
Canonical correlations are generalization of correlations to multi-dimensional variables.

@@

Multiple effect sizes measures based on canonical correlation are available.
\begin{itemize}
	\item Wilks' Lambda: $ \Lambda = \prod (1 - \phi_i^2) $
	\item Roy's Largest Root: $ \theta = \phi_{\max}^2 $
	\item Pillai's Trace: $ \tau = \sum \phi_i^2 $
\end{itemize}
We use Pillai's Trace in this paper. \todo[inline]{Why?}.

For the conditional case we use an approach similar to partial correlation. We
combine the mixed data residualization approach proposed in Ankan 2023. We then
use a canonical correlation based measure on these residuals to get an effect
size.

Benefits: Indifferent to one-hot encoding of the categorical variable.

\subsection{Connections to other measures}
This natually is a generalization of partial correlations. Furthermore, it is 
also a generalization of discrete case.

@@

>>>>

In this section, we develop a (conditional) measure of association for mixed
data. This measure is similar to partial correlation in the sense that it uses
a residualization approach and takes the association between the residuals.

Given two variables $ X $ and $ Y $, and a set of conditioanl variables $
\bm{Z} $, we are interested in computing the conditional association, $ \rho(X,
Y; Z) $. Similar to partial correlation, we start by computing the residuals
for $ X $ and $ Y $ using the $ \bm{Z} $ as the predictor varaibles and then
compute the assocation between the residuals. As the residuals for mixed
variable is not straightforward to define, we use the mixed data
residualization approach presented in Ankan and Textor 2023.

Given a variable $ X $ with predictor variables $ \bm{Z} $, we first train an
estimator, $ E_X $ for predicting $ X $ using $ \bm{Z} $. If the varialbes are
ordinal or categorical, the estimator needs to be probability estimators for
example a Random Forest Estimator or XGBoost. Given an estimator we can compute
the residuals as follows:
\begin{enumerate}
	\item If $ X $ is continuous, we simply take the difference between the true and estimated value.
		$$ R_X = X - E_X(\bm{Z}) $$
	\item If $ X $ is ordinal, we take the Li-Sheperd residual.
		$$ R_X = E_X(\bm{Z}) $$
	\item If $ X $ is categorical, we one hot encode the variable and Li-Sheperd residuals as well.
		$$ R_X = E_X(\bm{Z}) $$
\end{enumerate}
As a result of this residualization approach, the residuals are a matrix of
continuous values. The shape of this matrix depends on the type of variable. If
the variables are continuous or ordinal, we end up with a residual matrix of
size $ n \times 1 $. If the the variable is categorical we end up with a
residual matrix of size $ n \times (k-1) $, where $ k $ is the cardinality of
the categorical variables. We have dropped one of the one hot encoded columns
because it is redundant.

Next we need a measure to quantify the association between these residuals. One
of the general ways to have assocation between sets of variables is to use
canonical correlations. Canonical correlations are generalization of
correlations to multi-dimensional variables.

@@

Multiple measures of association based on canonical correlation are available such as:
\begin{itemize}
	\item Wilks' Lambda: $ \Lambda = \prod (1 - \phi_i^2) $
	\item Roy's Largest Root: $ \theta = \phi_{\max}^2 $
	\item Pillai's Trace: $ \tau = \sum \phi_i^2 $
\end{itemize}

We use Pillai's Trace in this paper. \todo[inline]{Why?}.

This measure of association has many of the desired properties:

\begin{enumerate}
	\item Is bounded between $ 0 $ and $ 1 $.
	\item Is independent of the sample size.
	\item Is indifferent to the one-hot encoding scheme used for categorical variables.
	\item In the case of both continuous variables, reduces to partial correlation.
\end{enumerate}

\subsection{Connections to other measures}
In the continuous variable case, this is equivalent to partial correlation.

This natually is a generalization of partial correlations. Furthermore, it is 
also a generalization of discrete case.

@@

Any two variables in a DAG that do not have an edge between them imply a
conditional independence test that is they are independent given the parents of
both the variables, $ X \ci Y | pa(X) \cup pa(Y) $. In the case of model
testing, a CI statement like this can be used to check whether the assumption
of the model holds in the data or is violated. For mixed data, a few different
CI tests have been proposed such as likelihood-ratio based, and residualization
based. However, if we just use the CI test to check the p-value we cannot use
it to find the modification that is able to improve the score the most.
